### 2022-05-03 response_new.tex

\begin{document}

\noindent Dear Editors:

We are re-submitting our manuscript, “The importance of the generation interval in investigating dynamics and control of new SARS-CoV-2 variants”, for consideration for publication in JRSI. This material has not previously been published elsewhere.

We had previously submitted the manuscript to JRSI (rsif-2021-0438) and were asked to revise with positive comments---for example, reviewer 3 said “The analysis is mathematically sound and offers a thorough investigation of an important concept that has been the subject of much discussion in the epidemiological community but, to my knowledge, hasn't been given the formal treatment it deserves until now.”

We revised the manuscript, addressing all comments, but the paper was rejected due to reviewer 2's comment that “On a second reading of this manuscript, I am much less optimistic about its value to infectious disease epidemiology.” We emailed the editor about this and were provided the opportunity to resubmit. We have now revised the manuscript to better highlight the novelty of our work.

The ongoing COVID-19 pandemic underlines the importance of characterizing the spread of new variants. Previous studies have particularly prioritized estimating the relative strength (i.e., the ratio of reproduction numbers) from relative speed (i.e., the differences between growth rates).

In this study, we challenge a common, sometimes implicit, assumption in estimating transmission advantage of new variants: that relative speed will remain constant.
We argue that constant relative strength is generally (though not always) more appropriate under changing conditions.
We also consider the possibility that new variants may have different generation intervals from the original variant and explore how such differences affect the dynamics and control of new variants. 
We discuss how neglecting these differences can lead to biased inference and provide a practical guidance for how such biases can be assessed in practice.
